---
output: github_document
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "##",
  fig.path = "man/images/"
)
set.seed(123)
```

# GMTM: Gaussian Mixture Topic Models

An R package for unsupervised or semi-supervised topic analysis of dense document vectors. **GMTM** performs clustering of document vectors using Gaussian mixture models. Document vectors can be source from any package but [wordvectors](https://github.com/koheiw/wordvector) works seamlessly. The underlying function is based on the [Armadillo library](https://arma.sourceforge.net/docs.html#gmm_diag) for fast computation. The algorithm is explain in the article below:

Sanderson, C., & Curtin, R. (2017). *An open source C++ implementation of multi-threaded Gaussian mixture models, k-means and expectation maximisation*. https://doi.org/10.1109/ICSPCS.2017.8270510


## Installation

From CRAN:

```{r, eval=FALSE}
#install.packages("GMTM") # not yet
```

From Github:

```{r, eval=FALSE}
devtools::install_github("koheiw/GMTM")
```

## Examples

**GMTM** is applied to identify topics of sentences from 20,000 news articles. Before clustering, [quanteda](https://github.com/quanteda/quanteda) and [wordvectors](https://github.com/koheiw/wordvector) are used for pre-processing textual data and training document vectors, respectively.

```{r, message=FALSE}
library(quanteda)
library(wordvector)
library(GMTM)

# pre-processing using quanteda
corp <- corpus_reshape(data_corpus_news2014)
toks <- tokens(corp, remove_punct = TRUE, remove_symbols = TRUE, remove_numbers = TRUE) |> 
  tokens_remove(stopwords("en"), min_nchar = 2)
dfmt <- dfm(toks)

# train document vectors using wordvector
wov <- textmodel_word2vec(toks, dim = 100)
dov <- as.textmodel_doc2vec(dfmt, wov)
```

### Unsupervised analysis

The basic usage of the package is discovering user-defined number of topics, `k`.

```{r}
gmm <- textmodel_gmm(dov, k = 10)
table(topics(gmm))
terms(gmm, data = dfmt)
```

### Semi-supervised analysis

It is also possible to use seed words to define topics: use `as.seedwords()` to create a seed matrix and pass it to `seeds`. 

```{r}
dict <- dictionary(list(economy = "econom*", politics = "politi*", 
                        security = c("securit*", "milita*"), 
                        sports = "sport*", crimes = "crime"))

seed <- as.seedwords(dict, wov, residual = 1)
sgmm <- textmodel_gmm(dov, seeds = seed)
table(topics(sgmm))
terms(sgmm, data = dfmt)
```

